---
title: "Map"
author: "M. Schrandt"
date: "7/16/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 3 Map

## 3.1 Read Spatial Data

### 3.11 Install packages

```{r}
# require() is like library() except returns FALSE if missing (vs error)
if (!require(librarian)){
  install.packages("librarian")
  library(librarian)
}
# librarian::shelf() is like library() except installs package if missing, 
#   even from Github if include owner/repo
shelf(
  NOAA-EDAB/ecodata,
  sf)
```

## 3.2 Get spatial data
```{r}
ecodata::epu_sf
```

```{r}
epu_sf <- ecodata::epu_sf %>% 
  st_transform(4326)
```

So we see a geometry list column.
```{r}
class(epu_sf)
```

```{r}
# "sf"         "data.frame"
g1 <- epu_sf$geometry[1]
# see in Environment pane, expand g1
```

```{r}
plot(epu_sf)
```

```{r}
plot(epu_sf["EPU"])
```

Where in the world is this? We can use `mapview` (a wrapper around `leaflet`) to locate on map.
```{r}
shelf(mapview)

mapview(epu_sf)
```

```{r}
shelf(leaflet)

leaflet() %>% 
  #addTiles() %>%
  addProviderTiles(providers$Esri.OceanBasemap) %>% 
  addPolygons(data = epu_sf)
```

## 3.3 Group by

sf is "tidy"

## 3.4 Extract from erddap
+ ERDDAP - Multi-scale Ultra-high Resolution (MUR) SST Analysis fv04.1, Global, 0.01°, 2002-present, Monthly - Data Access Form

+ CoastWatch ERDDAP: search for “SST”:

    + jplMURSST41mday: ERDDAP - Multi-scale Ultra-high Resolution (MUR) SST Analysis fv04.1, Global, 0.01°, 2002-present, Monthly - Data Access Form
    
```{r}
shelf(
  here,
  rerddap)

sst_gd_rds <- here("data/sst_gd.rds")

epu_bb <- st_bbox(epu_sf)
epu_bb
```

```{r}
sst_info <- info('jplMURSST41mday')
sst_info
```

```{r}
if (!file.exists(sst_gd_rds)){
  sst_gd <- griddap(
    sst_info,
    fields    = "sst",
    time      = c("2020-06-16", "2021-06-16"),
    longitude = epu_bb[c("xmin", "xmax")],
    latitude  = epu_bb[c("ymin", "ymax")])
  saveRDS(sst_gd, file = sst_gd_rds)  
}
sst_gd <- readRDS(sst_gd_rds)

sst_gd
```

```{r}
names(sst_gd)
```

```{r}
shelf(
  dplyr,
  ggplot2,
  mapdata)

# coastline
coast <- map_data(
  "worldHires", 
  xlim = epu_bb[c("xmin", "xmax")],
  ylim = epu_bb[c("ymin", "ymax")],
  lforce = "e")

sst_df_last <- sst_gd$data %>%
    filter(time == max(time))

# summary(sst_last)

ggplot(
  data = sst_df_last, 
  aes(x = lon, y = lat, fill = sst)) +
  geom_polygon(
    data = coast,
    aes(x = long, y = lat, group = group), fill = "grey80") +
  geom_tile() +
  scale_fill_gradientn(
    colors = rerddap::colors$temperature, na.value = NA) +
  theme_bw() + 
  ylab("Latitude") + 
  xlab("Longitude") +
  ggtitle("Latest SST")
```

```{r}
shelf(
  purrr,
  raster,
  sp,
  tidyr)
select <- dplyr::select

sst_tbl <- tibble(sst_gd$data) %>%
  mutate(
    # round b/c of uneven intervals
    #   unique(sst_gd$data$lon) %>% sort() %>% diff() %>% table()
    #     0.0099945068359375 0.0100021362304688
    lon  = round(lon, 2),
    lat  = round(lat, 2),
    date = as.Date(time, "%Y-%m-%dT00:00:00Z")) %>%
  select(-time) %>% 
  filter(!is.na(sst)) # 13M to 8.8M rows

sst_tbl_mo <- sst_tbl %>% 
  nest(data = c(lat, lon, sst)) %>% 
  mutate(
    raster = purrr::map(data, function(x) {
      #browser()
      sp::coordinates(x) <- ~ lon + lat
      sp::gridded(x)     <- T
      raster::raster(x)
    }))

sst_stk <- raster::stack(sst_tbl_mo$raster)
names(sst_stk) <- strftime(sst_tbl_mo$date, "sst_%Y.%m")
raster::crs(sst_stk) <- 4326
```

```{r}
shelf(stringr)

epu_sst_avg <- raster::extract(sst_stk, epu_sf, fun = mean, na.rm = T)
epu_sst_sd  <- raster::extract(sst_stk, epu_sf, fun = sd, na.rm = T)

epu_sst_tbl <- rbind(
  epu_sst_avg %>% 
    as_tibble() %>% 
    cbind(
      EPU  = epu_sf$EPU,
      stat = "mean") %>% 
    pivot_longer(-c(EPU, stat)),
  epu_sst_sd %>% 
    as_tibble() %>% 
    cbind(
      EPU = epu_sf$EPU,
      stat = "sd") %>% 
    pivot_longer(-c(EPU, stat))) %>% 
  mutate(
    EPU  = as.character(EPU),
    date = as.double(str_replace(name, "sst_", ""))) %>% 
  select(-name) %>% 
  pivot_wider(
    names_from  = EPU, 
    values_from = value)
```

```{r}
shelf(dygraphs)

epu_sst_tbl %>% 
  filter(stat == "mean") %>% 
  select(-stat) %>% 
  dygraph()
```

## Further Reading

R packages `mapview`, `leaflet`, `rerddap`